---
title: "STA130 R Data Wrangling and Summarizing Demo"
author: "Scott Schwartz"
date: "Fall 2022 Week 3"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, include=TRUE, echo=TRUE)
```


```{r}
library(tidyverse)
# install.packages("insuranceData")
library(insuranceData) 
data(AutoClaims)
glimpse(AutoClaims)
```

### Are these data tidy?

```{r}
AutoClaims %>% summary()
```

### How many `factor` levels do the `STATE` and `CLASS` have?

```{r}
# complete this
AutoClaims %>% group_by() %>% summarize() #%>% arrange()

#levels(AutoClaims$STATE)
#length(levels(AutoClaims$STATE))
#nlevels(AutoClaims$STATE)
#levels(AutoClaims$CLASS)
#length(levels(AutoClaims$CLASS))
#nlevels(AutoClaims$CLASS)
```

### Is the `summary()` output above confirmed using `dplyr` functions?

```{r}
# Complete this
AutoClaims %>% select() %>% filter() %>% 
  group_by() %>% 
  summarise(n=n(PAID), 
            mean=(PAID), median=(PAID), sd=(PAID), min=(PAID), max=(PAID),
            "25%th percentile"=(PAID), "75%th percentile"=(PAID))
```

### Is this data skewed? If so, is it right or left skewed?

```{r}
# Fix this
AutoClaims %>% ggplot("PAID") %>% geom_histogram()
```

### What's this code doing?

```{r}
AutoClaims %>% group_by(GENDER) %>% 
  summarise(n=n(), x0=sum(PAID>5000), x1=sum(PAID>5000)/n, x2=mean(PAID>5000), 
            x3=median(PAID>5000)) 
```

# Which of the following chunks of R code produces the most useful output for comparing the proportion of high insurance claims (over $10,000) for men and women?

```{r, fig.height=5, eval=FALSE}
# A
AutoClaims %>% ggplot(aes(x=PAID)) + 
  facet_wrap(~GENDER) + geom_histogram(bins=30, fill="gray", color="black")

# B
AutoClaims %>% ggplot(aes(x=GENDER, y=PAID)) + geom_boxplot()

# C
AutoClaims %>% group_by(GENDER) %>% 
  summarise(mean=mean(PAID), n=n(), new_var = sum(PAID > 10000)/n)
```


### Both of the following displays help us determine which state has the most insurance claims in our sample.  What is one advantage of each display?

```{r}
# A
AutoClaims %>% ggplot(aes(x=STATE)) + geom_bar() + coord_flip() 

# B
AutoClaims %>% group_by(STATE) %>% summarise(n=n())
```

### Do you like these better?

```{r}
# A
AutoClaims %>% ggplot(aes(x=reorder(STATE, STATE, function(x){length(x)}))) + geom_bar() + coord_flip() 
# https://stackoverflow.com/questions/5208679/order-bars-in-ggplot2-bar-graph

# B
AutoClaims %>% group_by(STATE) %>% summarise(n=n()) %>% 
  arrange(desc(n))
```


### Which of the following displays is most useful to explore whether there is an association between a driver's age and the amount of their insurance claim?

```{r}
# A
AutoClaims %>% ggplot(aes(x=AGE, y=PAID)) + geom_point(alpha=0.2)

# B
AutoClaims %>% group_by(AGE) %>% 
  summarise(mean=mean(PAID), med=median(PAID), n=n())

# C
AutoClaims %>% group_by(AGE) %>% 
  summarise(mean=mean(PAID), med=median(PAID), n=n()) %>% 
  ggplot(aes(x=AGE, y=mean)) + geom_point()

# D
AutoClaims %>% group_by(AGE) %>% 
  summarise(mean=mean(PAID), med=median(PAID), n=n()) %>% 
  ggplot(aes(x=AGE, y=mean, label=paste("n=",n,sep=""))) + geom_point() + 
  geom_text(vjust="bottom", hjust="left") 
# https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/paste
# https://ggplot2.tidyverse.org/reference/geom_text.html
```
